A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-01-30 ICLP Paper/samplepaper.tex
minor reformatting
section 3.2

\paragraph{Answer Set Programming (ASP)} \citep{lifschitz19a}
is a knowledge representation and reasoning paradigm geared for the effective 
modeling and solving of combinatorial (optimization) problems. % ASP has become an established paradigm for knowledge representation and reasoning. 
% ASP has proved its efficiency for solving the combinatorial optimization problem in different applications such as bioinformatics\cite{erdem2015generating,koponen2015optimizing}, databses\cite{caniupan2010consistency} and scheduling and industrial applications~\cite{dodaro2015allotment,dodaro2016combining,fabricius2020towards,dodaro2021operating}.
A (first-order) ASP program consists of \emph{rules} of the form
% \begin{lstlisting}[numbers=none,frame=none]
% h :- b#($_1$#), #(\dots#), b#($_n$#).
% \end{lstlisting}
\lstinline{h :- b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.},
in which the head \lstinline{h} is an atom
\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)} or a
choice
\lstinline{{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline|)}|
and each body literal \lstinline{b}$_i$ is an atom
\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)},
possibly preceded by the default negation connective \lstinline{not} and/or
followed by a condition
\lstinline|: c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$, 
a built-in comparison \lstinline{t}$_1\circ{}$\lstinline{t}$_2$
with $\circ\in\{\text{\lstinline{<}},\text{\lstinline{=}}\}$,
or an aggregate
\lstinline|t|$_0$\lstinline| = #count{t|$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$%
\lstinline| : c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$\lstinline|}|.
Each \lstinline|t|$_j$ denotes a term, i.e., a constant, variable, tuple,
or arithmetic expression, and each element \lstinline|c|$_k$ of a condition is
an atom that may be preceded by \lstinline{not} or a built-in comparison.
Roughly speaking, an ASP program is a shorthand for its ground instantiation,
obtainable by substituting variables with all of the available constants
and evaluating arithmetic expressions,
and the semantics is given by \emph{answer sets}, i.e.,
sets of (true) ground atoms such that all rules of the ground instantiation
are satisfied and allow for deriving each of the ground atoms in the head of
some rule whose body is satisfied.
The syntax of the considered ASP programs is a fragment of the modeling languages
described in \citep{cafageiakakrlemarisc20a,gehakalisc15a},
the ground instantiation process is detailed in \citep{kamsch21a},
and the answer set semantics is further elaborated in \citep{gehakalisc15a,lifschitz19a}.
%
% A logic program is a finite set of rules of the form 
% \begin{equation}
% 	a_0 \gets a_1, \ldots, a_m, \sim a_{m+1}, \dots, \sim a_n
% 	\label{rule1}
% \end{equation}
% where $a_i$ is an atom for $0 \leq i \leq n$ and ``$\sim$''  is a default negation. If $n$ quals to $0$, then a rule \eqref{rule1} is a fact. If $a_0$ is omitted, the rule is an integrity constraint. An \emph{atom} is an expression of the form $p(t_1,\ldots,t_l)$, where $p$ is a predicate and $t_1,\ldots,t_l$ are \emph{terms}. Each term can be a variable or a constant. A \emph{literal} is either an atom or its negation. 
% Given a rule $r$ of the form \eqref{rule1}, the set $H(r)=a_0$ denotes the \emph{head} atom and the set $B(r) = B^+(r) \cup B^-(r) = \{a_1,\dots,a_m\} \cup \{\sim a_{m+1}, \dots \sim a_n\}$ is the \emph{body} of $r$ where $B^+(r)$ and $B^+(r)$ represent the positive and negative body literals, respectively.

@@

In addition to J-EST decomposition, we have devised a similar ASP program for
\emph{Job-based Most Total Work Remaining (J-MTWR)} decomposition,
where the total order of operations is decreasing by the sum of processing
times for an operation and its successors in a job.
For example, we obtain the MTWR values~\lstinline{7}, \lstinline{12}, and~\lstinline{20}
for the first operation of job~\lstinline{1}, \lstinline{2}, or~\lstinline{3},
respectively, for the JSP instance in Listing~\ref{prg:facts},
matching the time for executing all three operations of each job.
Hence, the first operation of job~\lstinline{3} is considered as the most
important and is associated with the index~\lstinline{0}. The other
operations follow the total order taken for partitioning into time windows.
As J-MTWR values are decreasing along the sequence of operations in a job, the obtained time windows also respect the precedence of operations.

In addition to assessing decompositions in a purely Job-based fashion,
we have encoded \emph{Machine-based} decompositions \emph{M-EST} and \emph{M-MTRW} in which an operation from a bottleneck machine with the greatest sum of processing times for yet unordered operations is considered next.
For our example instance, the sum of processing times for operations 
is~\lstinline{12}, \lstinline{15}, or again \lstinline{12}, respectively,
for the machines~\lstinline{1}, \lstinline{2}, and~\lstinline{3}.
Hence, the operation to be inserted into the total order first is picked from
machine~\lstinline{2}, where the smallest EST or greatest MTWR value is
used for choosing a particular job/operation.
Unlike Job-based decomposition, this may lead to the choice of an operation
such that predecessor operations processed by other machines are yet unordered.
In this case, they are inserted into the total order directly before the
operation from a bottleneck machine,
and all machine loads are reduced by processing times
before determining the next machine to consider.

Moreover, we have devised dynamic versions of the Job- and Machine-based 
decomposition strategies in which partial schedules are taken into account,
and already scheduled operations from previous time windows may increase the
EST and MTWR values are used to order the yet unscheduled operations.
However, the principle of applying a stratified ASP program to determine
operations for the time window to schedule next is similar for all decomposition
strategies, and we experimentally evaluate them in Section~\ref{sec:experiments}.

@@

    % \\[-2.75mm]
    \botline% \hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
In Table~\ref{tab:Table02}, we compare other problem decomposition strategies
to the static Machine-based EST approach, providing the baseline abbreviated by M-EST.
Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR
variant, corresponding dynamic versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
determine EST and MTWR values w.r.t.\ partial schedules from previous time windows, as well as
static Job-based counterparts J-EST and J-MTWR in the first two rows.
For each decomposition strategy and JSP instance size, we present results for the number
of time windows,
which amounts to the next greater integer of the average number of interrupted optimization
runs in columns headed by Ints, leading to the shortest average makespan.
The considerably increased average makespans in the first two rows clearly indicate that
time windows determined with Job-based decomposition strategies are less adequate than
those investigating bottleneck machines in the first place, and then picking their operations
based on the smallest EST or greatest MTWR value as secondary criterion.
While Machine-based decomposition turns out to be advantageous,
there is no clear gap between the static M-EST and M-MTWR variants as well as their dynamic
versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$.
That is, the heuristic to schedule operations from highly loaded machines
as soon as possible works well for decomposing the considered JSP instances,
yet further solution quality differences due to minor deviations in the
decomposition strategy or optimization performance of \clingodl are negligible.

>>>>

\paragraph{Answer Set Programming (ASP)} \citep{lifschitz19a}
is a knowledge representation and reasoning paradigm geared for the effective 
modeling and solving of combinatorial (optimization) problems. % ASP has become an established paradigm for knowledge representation and reasoning. 
% ASP has proved its efficiency for solving the combinatorial optimization problem in different applications such as bioinformatics\cite{erdem2015generating,koponen2015optimizing}, databses\cite{caniupan2010consistency} and scheduling and industrial applications~\cite{dodaro2015allotment,dodaro2016combining,fabricius2020towards,dodaro2021operating}.
A (first-order) ASP program consists of \emph{rules} of the form
% \begin{lstlisting}[numbers=none,frame=none]
% h :- b#($_1$#), #(\dots#), b#($_n$#).
% \end{lstlisting}
\lstinline{h :- b}$_1$\lstinline{,}$\dots$\lstinline{,b}$_n$\lstinline{.},
in which the head \lstinline{h} is an atom
\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)} or a
choice
\lstinline{{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline|)}|
and each body literal \lstinline{b}$_i$ is an atom
\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)},
possibly preceded by the default negation connective \lstinline{not} and/or
followed by a condition
\lstinline|: c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$, 
a built-in comparison \lstinline{t}$_1\circ{}$\lstinline{t}$_2$
with $\circ\in\{\text{\lstinline{<}},\text{\lstinline{=}}\}$,
or an aggregate
\lstinline|t|$_0$\lstinline| = #count{t|$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$%
\lstinline| : c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$\lstinline|}|.
Each \lstinline|t|$_j$ denotes a term, i.e., a constant, variable, tuple,
or arithmetic expression, and each element \lstinline|c|$_k$ of a condition is
an atom that may be preceded by \lstinline{not} or a built-in comparison.
Roughly speaking, an ASP program is a shorthand for its ground instantiation,
obtainable by substituting variables with all of the available constants
and evaluating arithmetic expressions,
and the semantics is given by \emph{answer sets}, i.e.,
sets of (true) ground atoms such that all rules of the ground instantiation
are satisfied and allow for deriving each of the ground atoms in the head of
some rule whose body is satisfied.
The syntax of the considered ASP programs is a fragment of the modeling languages
described in \citep{cafageiakakrlemarisc20a,gehakalisc15a},
the ground instantiation process is detailed in \citep{kamsch21a},
and the answer set semantics is further elaborated in \citep{gehakalisc15a,lifschitz19a}.
%
% A logic program is a finite set of rules of the form 
% \begin{equation}
% 	a_0 \gets a_1, \ldots, a_m, \sim a_{m+1}, \dots, \sim a_n
% 	\label{rule1}
% \end{equation}
% where $a_i$ is an atom for $0 \leq i \leq n$ and ``$\sim$''  is a default negation. If $n$ quals to $0$, then a rule \eqref{rule1} is a fact. If $a_0$ is omitted, the rule is an integrity constraint. An \emph{atom} is an expression of the form $p(t_1,\ldots,t_l)$, where $p$ is a predicate and $t_1,\ldots,t_l$ are \emph{terms}. Each term can be a variable or a constant. A \emph{literal} is either an atom or its negation. 
% Given a rule $r$ of the form \eqref{rule1}, the set $H(r)=a_0$ denotes the \emph{head} atom and the set $B(r) = B^+(r) \cup B^-(r) = \{a_1,\dots,a_m\} \cup \{\sim a_{m+1}, \dots \sim a_n\}$ is the \emph{body} of $r$ where $B^+(r)$ and $B^+(r)$ represent the positive and negative body literals, respectively.

@@

In addition to J-EST decomposition, we have devised a similar ASP program for
\emph{Job-based Most Total Work Remaining (J-MTWR)} decomposition,
where the total order of operations is decreasing by the sum of processing
times for an operation and its successors in a job.
For example, we obtain the J-MTWR values~\lstinline{7}, \lstinline{12}, and~\lstinline{20}
for the first operation of job~\lstinline{1}, \lstinline{2}, or~\lstinline{3},
respectively, for the JSP instance in Listing~\ref{prg:facts},
matching the time for executing all three operations of each job.
Hence, the first operation of job~\lstinline{3} is considered as the most
important and is associated with the index~\lstinline{0}. The other
operations follow the total order taken for partitioning into time windows.
As J-MTWR values are decreasing along the sequence of operations in a job, the obtained time windows also respect the precedence of operations.

In addition to assessing decompositions in a purely Job-based fashion,
we have encoded \emph{Machine-based} decompositions \emph{M-EST} and \emph{M-MTRW} in which an operation from a bottleneck machine with the greatest sum of processing times for yet unordered operations is considered next.
For our example instance, the sum of processing times for operations 
is~\lstinline{12}, \lstinline{15}, or again \lstinline{12}, respectively,
for the machines~\lstinline{1}, \lstinline{2}, and~\lstinline{3}.
Hence, the operation to be inserted into the total order first is picked from
machine~\lstinline{2}, where the smallest M-EST or greatest M-MTWR value is
used for choosing a particular job/operation.
Unlike Job-based decompositions, this may lead to the choice of an operation
such that predecessor operations processed by other machines are yet unordered.
In this case, they are inserted into the total order directly before the
operation from a bottleneck machine, and all machine loads are reduced by processing times before determining the next machine to consider.

Moreover, we have devised \emph{dynamic versions} of the Job- and Machine-based 
decomposition strategies in which partial schedules are taken into account,
and already scheduled operations from previous time windows may change the EST and MTWR values used to order the yet unscheduled operations.
However, the principle of applying a stratified ASP program to determine
operations for the time window to schedule next is similar for all decomposition
strategies, and we experimentally evaluate them in Section~\ref{sec:experiments}.

@@

    % \\[-2.75mm]
    \botline% \hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
In Table~\ref{tab:Table02}, we compare other problem decomposition strategies
to the static Machine-based EST approach, providing the baseline abbreviated by M-EST.
Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR
variant, corresponding dynamic versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
determine EST and MTWR values w.r.t.\ partial schedules from previous time windows, as well as
static Job-based counterparts J-EST and J-MTWR in the first two rows.
For each decomposition strategy and JSP instance size, we present results for the number
of time windows,
which amounts to the next greater integer of the average number of interrupted optimization
runs in columns headed by Ints, leading to the shortest average makespan.
The considerably increased average makespans in the first two rows clearly indicate that
time windows determined with Job-based decomposition strategies are less adequate than
those investigating bottleneck machines in the first place, and then picking their operations
based on the smallest EST or greatest MTWR value as a secondary criterion.
While Machine-based decomposition turns out to be advantageous,
there is no clear gap between the static M-EST and M-MTWR variants as well as their dynamic
versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$.
That is, the heuristic to schedule operations from highly loaded machines
as soon as possible works well for decomposing the considered JSP instances,
yet further solution quality differences due to minor deviations in the
decomposition strategy or optimization performance of \clingodl are negligible.
